\thetitle

### 2025-05-13 workshop/PT.tex
almost 4 pages but needs work
TPG section + reorder

\section{Parallel tempering}
\label{sec:PT}
Pigeons.jl provides an implementation of distributed parallel tempering (PT) described 
in \cite{syed2021nrpt}, which we outline in Algorithm~\ref{alg:distributed_PT}.
This section gives both a brief overview of PT and some details of 
its distributed implementation.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=0.3\textwidth]{../img/bimodal.pdf}
    \caption{A simple bimodal distribution from which traditional 
    MCMC methods may struggle to obtain samples. Blue lines display output 
    from 1,000 iterations of a Metropolis-Hastings random walk MCMC algorithm.
    The sampler in this figure is visibly trapped in one of the two modes.}
    \label{fig:bimodal}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \begin{minipage}{0.15\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../img/heatmap_path_1.pdf}
      \caption*{$\pi_1$}
    \end{minipage}
    \begin{minipage}{0.15\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../img/heatmap_path_2.pdf}
      \caption*{$\pi_2$}
    \end{minipage}
    \begin{minipage}{0.15\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../img/heatmap_path_3.pdf}
      \caption*{$\pi_3$}
    \end{minipage}
    \begin{minipage}{0.15\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../img/heatmap_path_4.pdf}
      \caption*{$\pi_4$}
    \end{minipage}
    \begin{minipage}{0.15\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../img/heatmap_path_5.pdf}
      \caption*{$\pi_5$}
    \end{minipage}
    \begin{minipage}{0.15\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../img/heatmap_path_6.pdf}
      \caption*{$\pi_6$}
    \end{minipage}
    \caption{Heatmaps of six distributions lying on an annealing path 
    from a unimodal reference distribution  
    $\pi_1$, from which it is straightforward to obtain samples, and ending at 
    $\pi_N$, which is in this case the bimodal distribution from 
    Figure~\ref{fig:bimodal}. (Note that in this case the colours between the heatmaps 
    are not directly comparable because the densities of intermediate distributions 
    are not normalized.)}
    \label{fig:path}
\end{figure}

@@

@@

\begin{algorithm}[t]
\footnotesize
	\begin{algorithmic}[1]
    \Require Initial state $x_0$, sequence of distributions $\{\pi_i\}_{i=1}^N$, 
      number of iterations $T$, machine number $j$
    \State $\texttt{chain} \gets j$ \Comment{current \texttt{chain} number}
    \State $\texttt{dist\_array}[j] \gets j$ \Comment{\texttt{chain} $j$ is on machine $j$}

@@

\begin{figure*}[t]
    \centering
    \begin{minipage}{0.4\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../img/speedup_distributed.pdf}
    \end{minipage}
    \begin{minipage}{0.4\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../img/speedup_parallel.pdf}
    \end{minipage}
    \caption{
        Weak scaling results for PT in both the distributed (MPI, {\bf left}) 
        and parallel (Julia threads, {\bf right}) settings 
        with approximate 95\% confidence intervals for the speedup relative to single-thread execution.   
        The blue dotted line represents a hypothetical speedup equal to the 
        number of processes/threads. 
        The red dashed line indicates a speedup of one for reference.
    }
    \label{fig:PT_scaling}
\end{figure*}

>>>>

\section{Parallel tempering}
\label{sec:PT}

\begin{algorithm}[t]
	\begin{algorithmic}[1]
    \Require Initial state $x_0$, sequence of distributions $\{\pi_i\}_{i=1}^N$, 
      number of iterations $T$, machine number $j$
    \State $\texttt{chain} \gets j$ \Comment{current \texttt{chain} number}
    \State $\texttt{dist\_array}[j] \gets j$ \Comment{\texttt{chain} $j$ is on machine $j$}

Pigeons.jl provides an implementation of distributed parallel tempering (PT) described 
in \cite{syed2021nrpt}, which we outline in Algorithm~\ref{alg:distributed_PT}.
This section gives both a brief overview of PT and some details of 
its distributed implementation.

\begin{figure}[t]
    \centering
    \includegraphics[width=0.25\textwidth]{../img/bimodal.pdf}
    \caption{A simple bimodal distribution from which traditional 
    MCMC methods may struggle to obtain samples. Blue lines display output 
    from 1,000 iterations of a Metropolis-Hastings random walk MCMC algorithm.
    The sampler in this figure is visibly trapped in one of the two modes.}
    \label{fig:bimodal}
\end{figure}

@@

@@

@@

% NB: this is placed here just to make it appear in a different page
\begin{figure}[t]
    \centering
    \begin{minipage}{0.15\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../img/heatmap_path_1.pdf}
      \caption*{$\pi_1$}
    \end{minipage}
    \begin{minipage}{0.15\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../img/heatmap_path_2.pdf}
      \caption*{$\pi_2$}
    \end{minipage}
    \begin{minipage}{0.15\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../img/heatmap_path_3.pdf}
      \caption*{$\pi_3$}
    \end{minipage}
    \begin{minipage}{0.15\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../img/heatmap_path_4.pdf}
      \caption*{$\pi_4$}
    \end{minipage}
    \begin{minipage}{0.15\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../img/heatmap_path_5.pdf}
      \caption*{$\pi_5$}
    \end{minipage}
    \begin{minipage}{0.15\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../img/heatmap_path_6.pdf}
      \caption*{$\pi_6$}
    \end{minipage}
    \caption{Heatmaps of six distributions lying on an annealing path 
    from a unimodal reference distribution  
    $\pi_1$, from which it is straightforward to obtain samples, and ending at 
    $\pi_N$, which is in this case the bimodal distribution from 
    Figure~\ref{fig:bimodal}. (Note that in this case the colours between the heatmaps 
    are not directly comparable because the densities of intermediate distributions 
    are not normalized.)}
    \label{fig:path}
\end{figure}

@@

@@

\begin{figure*}[t]
    \centering
    \includegraphics[width=0.33\textwidth]{../img/speedup_distributed.pdf}
    \hspace{4em}
    \includegraphics[width=0.33\textwidth]{../img/speedup_parallel.pdf}
    \caption{
        Weak scaling results for PT in both the distributed (MPI, {\bf left}) 
        and parallel (Julia threads, {\bf right}) settings 
        with approximate 95\% confidence intervals for the speedup relative to single-thread execution.   
        The blue dotted line represents a hypothetical speedup equal to the 
        number of processes/threads. 
        The red dashed line indicates a speedup of one for reference.
    }
    \label{fig:PT_scaling}
\end{figure*}
